---
title: "Untitled"
author: "Katja Berghaus"
date: "21 Mai 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r Libraries}
library(dplyr)
library(tidyr)
```
Read in data from https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1

```{r}
catch_df <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                    stringsAsFactors = FALSE)
head(catch_df)
```
```{r}
unique(catch_df$notesRegCode)
```
```{r}
catch_df <- catch_df %>% 
  select(-All, -notesRegCode) %>% 
  gather(key = Species, value = Catch, -Region, -Year) %>% 
  rename(catch_thousands = Catch)
  
```

Found a typo in original data, corrected it, converted `catch_thousands`column to numeric, and created a new column with units# of fish
#Gather

key (column names) and value(values in the cells)


-Region, -Year -> don't gather Region and Year, alternatively I could've called all the columns that we wanted in it


Syntax of rename: Old name = new name!
which(is.na)== TRUE -> which one is NA in my data set
Now change I into a 1 and then go on with what we wanted to do initially
```{r}
catch_df<- catch_df %>% 
  mutate(catch_thousands = ifelse(catch_thousands == "I", 1, catch_thousands)) %>% 
  mutate(catch_thousands = as.integer(catch_thousands)) %>% 
  mutate(catch = catch_thousands *1000)

head(catch_df)
```

#Analysis

```{r}
catch_df %>% 
  group_by(Species, Year) %>% 
  summarize(mean_catch = mean(catch), num_obs = n())
```
```{r}
catch_df %>% 
  filter(Species == "Chinook") %>% 
  group_by (Region) %>% 
  summarise(mean_catch = mean(catch)) %>% 
  arrange(desc(mean_catch)) #sort in descending order by catch
```

